---
title: "Shopify Technical Challenge"
author: "Silvane Paixao (silpai)"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: Cayman
    highlight: github
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load Packages & libraries

if(!require(tidyverse)) install.packages("tidyverse", repos="http://cran.us.r-project.org")
if(!require(gsheet)) install.packages("gsheet", repos="http://cran.us.r-project.org")                    # load google sheet
if(!require(lubridate)) install.packages("lubridate", repos="http://cran.us.r-project.org")              # format date
if(!require(ggthemes)) install.packages("ggthemes", repos="http://cran.us.r-project.org") 
if(!require(formattable)) install.packages("formattable", repos = "http://cran.us.r-project.org")        # To format tables
if(!require(htmltools)) install.packages("htmltools", repos = "http://cran.us.r-project.org")            # To format tables
if(!require(DT)) install.packages("DT", repos = "http://cran.us.r-project.org")            # To format tables
if(!require(prettydoc)) install.packages("prettydoc", repos = "http://cran.us.r-project.org")            # To format R Markdown




library(tidyverse)
library(gsheet)
library(lubridate)
library(ggthemes)
library(formattable)
library(htmltools)
library(DT)
library(prettydoc)




#update.packages ()

options(digits=2)
options(scipen=999) 
############################################  Load Data  ############################################
url<-"https://docs.google.com/spreadsheets/d/16i38oonuX1y1g7C_UAmiK9GkY7cS-64DfiDMNiR41LM/edit#gid=0"
data<-gsheet2tbl(url) # retrieve data from googlesheet

#str(data)

data$created_date <- as.Date(data$created_at)  # convert str to date
data<-data %>% mutate(year=year(created_date), month=month(created_date, label=TRUE),day=day(created_date))


formattable (data)
```

## Question 1

The Average Order Value (AOV) is the average $ spent each time a customer places an order. It is calculated using the total revenue by the total number of customers' orders. 
Validating the AOV stated in the challenge instructions for the existing 30 days, it was confirmed that the monthly AOV is $3145. 

**Table 1** shows the daily AOV, delta = daily AOV - monthly AOV and its delta classification. 

```{r , echo = FALSE}
TAB_montly_AOV<- data %>% 
  arrange(desc(created_date)) %>%
  group_by(created_date) %>% 
  summarise(order_day=n(), order_amount_day=sum(order_amount)) %>%
  mutate (daily_AOV = order_amount_day/order_day, AOV30d=sum(order_amount_day)/sum(order_day), 
          delta=daily_AOV - AOV30d, margin = ifelse(daily_AOV>=AOV30d,  "above monthly AOV", "below monthly AOV"))  


formattable (TAB_montly_AOV, align= c("c","c","c","c","c","c","l"),
             list ('delta'=formatter ("span",
                                      style=x~style (font.weight="bold",
                                                     color= ifelse (x>0,"#71CA97", ifelse (x<0, "#ff7f7f", "black")))
             )))

AOV_30d<- sum(TAB_montly_AOV$order_amount_day)/sum(TAB_montly_AOV$order_day)


```


Considering the margin data from **Table 1**, it can be seen at **Table 2** that 40% of the days had revenue above of the the monthly AOV ($3145).


```{r , message=FALSE, warning=FALSE, echo=FALSE}
TAB_margin<-transform(as.data.frame(table(TAB_montly_AOV$margin)))
TAB_margin$Prop <- prop.table(table(TAB_montly_AOV$margin))
formattable(TAB_margin)
```


**Daily AOV** graphic shows that only 4 days in March had daily AOV above $7500 or 2x monthly AOV; one of the days had a daily AOV outlier of $12500 or almost 4x monthly AOV. 


```{r,  message=FALSE, warning=FALSE}
dailyAOV<- TAB_montly_AOV %>% 
  ggplot(aes(x=created_date, y=daily_AOV))+
  geom_point(col='#014d64')+
  labs(x='', y='daily AOV')+
  theme_economist()+
  labs(title = 'Daily AOV',
       subtitle= "Average Order Value",
       caption='Reference line: monthly AOV'
  )+
  geom_hline(yintercept=AOV_30d,color='#e5001f',lwd=0.4)+
  theme(plot.title = element_text(family = 'Avenir Next Condensed', hjust=0,size=18,margin=margin(-7,0,10,0)),
        plot.subtitle = element_text(family = 'Avenir Next Condensed',size=12, margin=margin(-5,-1,3,0)),
        plot.caption = element_text(family = 'Avenir Next Condensed', size=9, hjust = 0,margin=margin(34,0,-45,0)),  
        axis.text = element_text(face='bold'),
        axis.title.x = element_text(family = 'Avenir Next Condensed',face = 'bold', margin=margin(12,0,-12,0)),
        axis.title.y = element_text(family = 'Avenir Next Condensed',face = 'bold', margin=margin(0,12,0,0)),
        plot.margin = margin(1,1.5,2,0.5, 'cm'))
dailyAOV



```


Comparing *Daily AOV - Monthly AOV*, the maximum daily margin gain was $9226 and the maximum daily margin lost was $-2867.

```{r, echo=FALSE}
summary(TAB_montly_AOV$delta)

```


```{r echo=TRUE,  message=FALSE, warning=FALSE}
daily_margin<- TAB_montly_AOV %>% 
  ggplot(aes(x=created_date,y=delta))+
  geom_bar(stat="identity", position="identity", aes(fill=margin)) +
  labs(x='', y="delta")+
  theme_economist()+
  labs(title = 'AOV Margins',
       subtitle= "Daily AOV - Monthly AOV",
       caption='Source: Spotify'
  )+
  
  scale_fill_manual(values=c("steelblue","firebrick1")) +
  
  theme(plot.title = element_text(family = 'Avenir Next Condensed', hjust=0,size=18,margin=margin(-7,0,10,0)),
        plot.subtitle = element_text(family = 'Avenir Next Condensed',size=12, margin=margin(-5,-1,3,0)),
        plot.caption = element_text(family = 'Avenir Next Condensed', size=9, hjust = 0,margin=margin(34,0,-45,0)),  
        axis.text = element_text(face='bold'),
        axis.title.x = element_text(family = 'Avenir Next Condensed',face = 'bold', margin=margin(12,0,-12,0)),
        axis.title.y = element_text(family = 'Avenir Next Condensed',face = 'bold', margin=margin(0,12,0,0)),
        plot.margin = margin(1,1.5,2,0.5, 'cm')) + 
  ylim(-5000,10000)
daily_margin

```                
             



**Loyalty Analysis - Metric to report**

Calculation of the **Purchase Frequency** indicated that a customer purchased 16.6 times in march (see **Table 3**:*number of orders = 5000*, *number of unique user=301*).

```{r echo=FALSE,  message=FALSE, warning=FALSE}
Purchase_Freq <- data%>% arrange(desc(user_id))%>%
  group_by(month)%>%  summarise(orders_count=length(order_id), user_count=length(unique(user_id)),Purchase_Freq = orders_count/ user_count)
formattable (Purchase_Freq)

```

Looking at the data, I was curious to know if customers were loyal to a particular store. **Table 4** shows the number of times a user ordered at the same store and AOV per user per store.

Customers shopped in multiple stores (except for user 607, who only shopped at store 42 for 17 times). 
```{r echo=FALSE,include=FALSE,  message=FALSE, warning=FALSE}

TAB_AOV_user_per_shop<-data%>% arrange(desc(user_id))%>%
  group_by(user_id,shop_id)%>% 
  summarise(orders_count=n(), order_amount_sum=sum(order_amount), AOV_user_shop= sum(order_amount_sum)/sum(orders_count)) %>%
  mutate(user_shop_class=ifelse (AOV_user_shop < AOV_30d, "C: less than monthly AOV",                               
                            ifelse(AOV_user_shop  >= AOV_30d & AOV_user_shop <= AOV_30d*2, "B: >= monthly AOV <= 2x Monthly AOV", "A: Above 2x Monthly AOV")))

#formattable (TAB_AOV_user_per_shop, align= c("c","c","c","c","c","l"),
            # list ('AOV_user_shop'=formatter ("span",
                                             # style=x~style (font.weight="bold",
                                                           #  color= ifelse (x<AOV_30d,"#ff7f7f", ifelse (x >= AOV_30d & x <= AOV_30d*2, "yellow", "#71CA97")))
           #  )))

```


```{r , echo=FALSE}
DT::datatable(TAB_AOV_user_per_shop, filter='top')
```



Considering that each user shopped in multiple locations, **Table 5**  shows that 99% AOV user per store  (4551 stores) were classified as class C: less than monthly AOV, and the 1% (46 stores) were class A: Above 2x Monthly AOV.


```{r, echo=FALSE}
# Compute percentages
con_AOV_user_per_shop<-transform(as.data.frame(table(TAB_AOV_user_per_shop$user_shop_class)))
con_AOV_user_per_shop$Prop <- prop.table(table(TAB_AOV_user_per_shop$user_shop_class))
formattable(con_AOV_user_per_shop) 
```



```{r , echo=FALSE,  message=FALSE, warning=FALSE, out.width = "150%"}


con_AOV_user_per_shop %>% ggplot(aes(x = Var1, weight = Prop)) +
  geom_bar() + 
  theme_economist() +
  labs(title = 'User Class Analysis',
       subtitle= "User Class % per each store",
       caption='Source: Spotify', 
       x='', y='')+
  theme(plot.title = element_text(family = 'Avenir Next Condensed', hjust=0,size=18,margin=margin(-7,0,10,0)),
        plot.subtitle = element_text(family = 'Avenir Next Condensed',size=12, margin=margin(-5,-1,3,0)),
        plot.caption = element_text(family = 'Avenir Next Condensed', size=9, hjust = 0,margin=margin(34,0,-45,0)),  
        axis.text = element_text(face='bold'),
        axis.title.x = element_text(family = 'Avenir Next Condensed',face = 'bold', margin=margin(12,0,-12,0)),
        axis.title.y = element_text(family = 'Avenir Next Condensed',face = 'bold', margin=margin(0,12,0,0)),
        plot.margin = margin(1,1.5,2,0.5, 'cm'))+
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 10))


```

**Table 6** displays the number of stores that each user ordered. 


```{r echo=FALSE,  message=FALSE, warning=FALSE}
TAB_AOV_user_ALL_shop<-data%>% 
  group_by(user_id)%>% 
  summarise(shop_id_count=length(unique(shop_id)), order_amount_ALL_sum=sum(order_amount), AOV_user_ALL_shop= sum(order_amount_ALL_sum)/sum(shop_id_count)) %>%
  arrange((shop_id_count))   %>%
  mutate( user_ALL_shop_class=ifelse (AOV_user_ALL_shop < AOV_30d, "C: less than monthly AOV",                               
                                 ifelse(AOV_user_ALL_shop  >= AOV_30d & AOV_user_ALL_shop <= AOV_30d*2, "B: >= monthly AOV <= 2x Monthly AOV", "A: Above 2x Monthly AOV ")))


#formattable (TAB_AOV_user_ALL_shop, align= c("c","c","c","c","l"),
            # list ('AOV_user_ALL_shop'=formatter ("span",
                                             # style=x~style (font.weight="bold",
                                                             color= ifelse #(x<AOV_30d,"red", ifelse (x >= AOV_30d & x <= AOV_30d*2, "orange", "green")))

        #)))

                                                             DT::datatable(TAB_AOV_user_ALL_shop, filter='top')

```


Considering the 301 unique users, Table 7 indicates that the majority of individual customers (92%) have been being classified as C: less than monthly AOV, followed by 7%  who spent between the monthly AOV and 2x monthly AOV. Only 1% of the customers were A: above 2x monthly AOV; they should be treated as VIP.


```{r, echo=FALSE}
con_AOV_user_per_ALL_shop<-transform(as.data.frame(table(TAB_AOV_user_ALL_shop$user_ALL_shop_class)))
con_AOV_user_per_ALL_shop$Prop <- prop.table(table(TAB_AOV_user_ALL_shop$user_ALL_shop_class))
formattable(con_AOV_user_per_ALL_shop)
```


```{r echo=FALSE,  message=FALSE, warning=FALSE}


con_AOV_user_per_ALL_shop %>% ggplot(aes(x = Var1, weight = Prop)) +
  geom_bar() + 
  theme_economist() +
  labs(title = 'User Class Analysis',
       subtitle= "User Class % per overall stores",
       caption='Source: Spotify', 
       x='', y='')+
  theme(plot.title = element_text(family = 'Avenir Next Condensed', hjust=0,size=18,margin=margin(-7,0,10,0)),
        plot.subtitle = element_text(family = 'Avenir Next Condensed',size=12, margin=margin(-5,-1,3,0)),
        plot.caption = element_text(family = 'Avenir Next Condensed', size=9, hjust = 0,margin=margin(34,0,-45,0)),  
        axis.text = element_text(face='bold'),
        axis.title.x = element_text(family = 'Avenir Next Condensed',face = 'bold',  margin=margin(12,0,-12,0)),
        axis.title.y = element_text(family = 'Avenir Next Condensed',face = 'bold', margin=margin(0,12,0,0)),
        plot.margin = margin(1,1.5,2,0.5, 'cm')) +
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 10)) +
  ylim (0,1)

```


## Question 2 - SQL ## 

**1) How many orders were shipped by Speedy Express in total?**

SELECT sh.ShipperName , COUNT(sh.ShipperName) AS ShipperCount <br />
FROM Orders o <br />
INNER JOIN Shippers sh ON <br />
sh.ShipperID= o.ShipperID <br />
WHERE sh.ShipperName='Speedy Express';


|ShipperName    |	ShipperCount |
|:---           |:---          |
|Speedy Express |	54           |


**2) What is the last name of the employee with the most orders?**

SELECT e.LastName, COUNT(od.OrderID) AS OrderCount <br />
FROM Orders o <br />
INNER JOIN Employees e ON <br />
e.EmployeeID= o.EmployeeID <br />
INNER JOIN OrderDetails od ON <br />
od.OrderID= o.OrderID <br />
GROUP BY e.LastName <br />
ORDER BY OrderCount DESC <br />
LIMIT 3;


|LastName |	OrderCount |
|:---           |:---          |
|Peacock |	123|
|Davolio|	78|
|Leverling|	74|


**3) What product was ordered the most by customers in Germany?**
  
SELECT p.ProductName, cus.Country, od.Quantity <br />
FROM Orders o <br />
INNER JOIN OrderDetails od ON <br />
od.OrderID= o.OrderID <br />
INNER JOIN Products p ON <br />
p.ProductID= od.ProductID <br />
INNER JOIN Customers cus ON <br />
cus.CustomerID= o.CustomerID <br />
WHERE Country = 'Germany' <br />
GROUP BY p.ProductName <br />
order by Quantity DESC <br />
LIMIT 3;



|ProductName	|Country |	Quantity |
|:---           |:---          |:---          |
|Steeleye Stout |	Germany	|100 |
|Teatime Chocolate Biscuits	|Germany|	80|
|Raclette Courdavault	|Germany |	70|


